The human in the loop: volunteer computing as a socio-technical system

### 2016-03-01 Alife16/paper.tex
Final changes after submission
Closes #100
Closes #102

\begin{abstract}
Volunteer computing is a form of distributed computing where users
decide on their participation and the amount of time and other resources they will
``lend''. This makes them an essential part of the algorithm and of
the performance of the whole system. As a sociotechnical system, this
participation follows some patterns and in this paper we examine the
result of several volunteer distributed evolutionary computation
experiments and try to find out which are those patterns and what
makes an experiment successful or not, including the feedback loop
that is created between the users and the algorithm itself. In this
paper we will examine what kind of patterns we can expect and what
influence those patterns have on the implementation of the
evolutionary algorithm. 
%Pablo: The patterns we are looking for are... 
\end{abstract}

@@

This loose connection provides a low-overhead way to connect desktop
experiments with volunteer-based ones, with all of them contributing
to the pool, but every one of them working as separate island 
carrying out their own evolutionary algorithms. This is why NodIO is
proposed mainly as a {\em complement} to existing resources such as
desktop systems or laptops. As long as it provides a non-null
computational capability that can help existing resources find the
solution faster it will have found its purpose. The main use case is
someone setting up NodIO in the cloud, writing a fitness evaluation
function and running a client from his or her own computer, but
requesting help in social networks for additional resources. That is
why, in this system,  we have considered the whole {\em social} aspect
in the design, with issues related to security, trust and privacy
among others. The computing system becomes a {\em techno-social
%Pablo: In abstract is named 'sociotechnical'. Which one is the correct form for the rest of the paper?
  system} \citep{vespignani2009predicting}. In this paper, we are going
to focus on measuring the response of users to experiments, but at the
%Pablo: which metric are we going to measure this response?
same time we will also focus on the technical aspects of the server and how
those might change the behavior of users, improving the capability of % Mario: How these might change
the system.

Our research group is committed to open science, and we think this is
a very important part of the techno-social system. By being
transparent, incentives to cheat are reduced and, in fact, we have
detected no issue for the time being. Next we present the state of the art in web-based
volunteer-based systems along with attempts to predict and model its
behavior.

>>>>

\begin{abstract}
Volunteer computing is a form of distributed computing where users
decide on their participation and the amount of time and other
resources they will ``lend''. This makes them an essential part of the
algorithm and of the performance of the whole system. As a
sociotechnical system, this participation follows some patterns and in
this paper we examine the result of several volunteer distributed
evolutionary computation experiments and try to find out which are
those patterns and what makes an experiment successful or not,
including the feedback loop that is created between the users and the
algorithm itself. In this paper we will examine what kind of patterns
we can expect and what influence those patterns have on the
implementation of the evolutionary algorithm.
\end{abstract}

@@

This loose connection provides a low-overhead way to connect desktop
experiments with volunteer-based ones, with all of them contributing
to the pool, but every one of them working as separate island carrying
out their own evolutionary algorithms. This is why NodIO is proposed
mainly as a {\em complement} to existing resources such as desktop
systems or laptops. As long as it provides a non-null computational
capability that can help existing resources find the solution faster
it will have found its purpose. The main use case is someone setting
up NodIO in the cloud, writing a fitness evaluation function and
running a client from his or her own computer, but requesting help in
social networks for additional resources. That is why, in this system,
we have considered the whole {\em social} aspect in the design, with
issues related to security, trust and privacy among others. The
computing system becomes a {\em sociotechnical} system}
\citep{vespignani2009predicting}. In this paper, we are going to focus
on measuring the response of users to experiments, that is, the time
they spend running it, but at the same time we will also focus on the
technical aspects of the server and how these might change the
behavior of users, improving the capability of the system.

Our research group is committed to open science, and we think this is
a very important part of the techno-social system. By being
transparent, incentives to cheat are reduced and, in fact, we have
detected no issue for the time being. Next we present the state of the
art in web-based volunteer computing systems along with attempts to
predict and model its behavior.
